---
title: "Penguins size comparisons"
author: "Morganne Sigismonti"
date: "11/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
library(effsize)
library(broom)
library(ggbeeswarm)
library(car)
```

# Part 1. T-tests

Mass comparison: makle and female chinstrap penguins
```{r}
chinstraps <- penguins %>% 
  filter(species == "Chinstrap")

#Histrogram
ggplot(data = chinstraps, aes(x = body_mass_g)) +
  geom_histogram(bins = 10) +
  facet_wrap(~sex)

## QQ plot
# to compare our sample to data if it were perfectly drawn from a normal distribution

ggplot(data = chinstraps, aes(sample = body_mass_g)) +
  geom_qq() +
  facet_wrap(~sex)

### Really close to normal distribution! As close as we can probably get!

## Get descriptive statistrics

chinstraps %>%
  group_by(sex) %>% 
  summarize(
    mean_mass = mean(body_mass_g),
    median_mass = median(body_mass_g),
    sd_mass = sd(body_mass_g),
    sample_size = n()
  )
```

Get vectors of body_mass_g for female and male chinstrap penguins separately.

```{r}
chinstraps_f <- chinstraps %>% 
  filter(sex == "female") %>% 
  pull(body_mass_g)

chinstraps_m <- chinstraps %>% 
  filter(sex == "male") %>% 
  pull(body_mass_g)
```


### T-tests, a couple diffferent ways

1. is there a significant difference in body mass between male and female chinstrap penguins?

```{r}
chinstraps_ttest <- t.test(chinstraps_f, chinstraps_m)
chinstraps_ttest

cohen.d(chinstraps_f, chinstraps_m)
```

null: means are equal for male and femal chinstrap penguins
p value says: super low chance that we ciuld have taken these two samples and gotten measn that are at least as different as ours are if they're drawn from populations with the same mean sufficient evidence to reject the null hypothesis.  at least as different as ours are by random chance if they were drawn from pops with the same mean

large effect size: pretty clear difference between the groups

There is a significant difference in body mass, and a large effect size (Cohen's *d* = -1.26). The way I'm *writing this* is not reproducible. How can we use in-line referencing of outcomes so that the values would automatically update if the values did? 

Use `broom::tidy()` to get messy model outputs into tidy data frames that are easy to reference! 

```{r}
chinstraps_out <- broom::tidy(chinstraps_ttest)
chinstraps_out
```


2. Are male chinstrap penguins significantly larger than female chinstrap penguins? 

This is a *one-sided* test - we are only interested in the upper tail. We will still use the `t.test()` function, but update the default `alternative = ` argument from "two-sided" to "greater" or "less."

```{r}
# "Is mean male mass significantly GREATER than female mass?"
t.test(chinstraps_m, chinstraps_f, alternative = "greater")

# Yields the same statistical outcomes as "Is mean female mass significantly LESS than male mass?" 
t.test(chinstraps_f, chinstraps_m, alternative = "less")
```
reject the null hypothesis and retain the alt hyp

male chinstrap penguins body mass is significantly greater than female chinstrap penguins body mass

## One-way ANOVA

## 1. Table of counts by species and sex

```{r}
penguin_counts <-penguins %>% 
  count(species, sex)

penguin_counts
```

Create a subset of only female penguins
```{r}
penguin_f <- penguins %>% 
  filter(sex == "female")

ggplot(data = penguin_f, aes(x = flipper_length_mm)) +
  geom_histogram(aes(fill = species)) +
  facet_wrap(~species, scales = "free")
## scales = free lets them each have their own x and y axises. looking at stuff between groups

ggplot(data = penguin_f, aes(sample = flipper_length_mm)) +
  geom_qq(aes(color = species)) +
  facet_wrap(~species, scales = "free")

## super linear! probs drawn from normnally distributed populations
```

### Summary statistics

Starting from penguins_f, create a summary table that groups by species and returns the following:

- mean of flipper length
- sd""
- sample size
- variance 'var'
- standard error

```{r}
penguin_f_summary <- penguin_f %>%
  group_by(species) %>% 
  summarize(
    mean_flipper_length = mean(flipper_length_mm),
    sd_flipper_length = sd(flipper_length_mm),
    sample_size = n(),
    se_flipper_length = sd(flipper_length_mm)/sqrt(n()),
    var_flipper_length = var(flipper_length_mm)
  )

leveneTest(flipper_length_mm ~ species, data = penguin_f)

## retain the null hypothesis looking at Pr(>F)


```


### Beeswarm plot with mean += 1 sd
- shows the distribution of my observations
where the greater densities of populationa are

```{r}
ggplot() +
  geom_beeswarm(data = penguin_f,
                aes(x = species, y = flipper_length_mm),
                color = "gray70",
                size = 3,
                alpha = .5) +
  geom_point(data = penguin_f_summary,
             aes(x = species,
                 y = mean_flipper_length),
                 size = 4) +
  geom_errorbar(data = penguin_f_summary,
                aes(x = species,
                    ymax = mean_flipper_length + sd_flipper_length,
                    ymin = mean_flipper_length - sd_flipper_length),
                width = 0.1) +
  theme_light()

```
### One-way ANOVA

```{r}
penguin_aov <- aov(flipper_length_mm ~ species, data = penguin_f)
summary(penguin_aov)

TukeyHSD(penguin_aov)
```
### F stat = ratio of between group variance / within group variance
large the variance between between groups is large compared to wihtin groupas
negligible small prob that we culd have gotten sample means that are at least this different if all three of our samples were actually drawn from populations with the same mean flipper length

We would reject our null hypothesis that the means for all of these groups, for chin strap adelli and gentoo penguins, are the same

Post hoc test tells us that there is a difference between all penguin groups 


Type III anova with the car package:
- if you don't have a very balanced design (otherwise tpye 1 and 2 work well)

```{r}
penguin_lm <- lm(flipper_length_mm ~ species, data = penguin_f)
penguin_aov_t3 <- Anova(penguin_lm, type = "III")
penguin_aov_t3
```

